Marine species do not need visas: the transboundary nature of the world’s exploited marine species

### 2020-03-04 Manuscript_Draft.Rmd
Submitted manuscript with figures and references

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada, ^2^Yale University, New Haven, Connecticut, USA, ^3^Stockholm Resilience Centre, Stockholm University, Stockholm, Sweden"
# subtitle: "[DRAFT DO NOT SHARE]"
subtitle: "Most neighbour nations share marine species"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ./References/FishForVisa_Ref.bib
csl: ./References/science.csl
---

@@

# One Sentence Summary (125/~125 characters)
<!-- *A brief teaser statement highlighting main result of the paper, understandable by a scientist not in your field, without jargon or abbreviations. This will appear online adjacent to the title and should not repeat phrases already present there* -->

@@

# Abstract (150/~125 words)
<!-- *An opening sentence that sets the question that you address and is comprehensible to the general reader, background content specific to this study, results, and a concluding sentence. It should be a single paragraph.* -->

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighbouring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 80% of the global-EEZs catches (47 million tonnes) and 78.5% of fisheries revenue (USD 65 billion) in last decade. Some countries received over 70% of their economic benefits from a few shared resources. This study highlights the large and under-recognized challenge in achieving international fisheries management goals, particularly as species redistribute under climate change.

# Main Article (3200/~2500 words)
<!-- # Introduction (~555/500 words) -->

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors defined by their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. Thus, a species can have a broader distribution (cosmopolitan) or be endemic [@Nelson:2016kk]. However, fisheries management is predicated on the definition of “stocks” that, in most cases, are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, established political boundaries across the distribution of many species, creating shared species between nations. Theories and empirical experience have show that fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@Munro:2002uf; @Bailey:2010ga]. For instance, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and for fisheries management [@Miller:2013iv].

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that are shared by neighboring nations (hereafter called transboundary species [@Munro:2002uf]) is still unclear. Early data limitations on species' distributions allowed only for an informed guess on the number of transboundary species [@Caddy:1997ue] while a more recent study explored the historical contribution of shared stocks following a litterature review [@Teh:2015gd]. Thus, no method has included the actual distribution of species in the quantification of transboundary stocks. Here, we aimed to estimate the number of exploited marine species shared by neighbouring countries and determine their contribution to global and regional catch and fisheries revenue. Moreover, we categorize species according to their exploitation rate while identifying differences by habitat preference. We hipothesize that the methodological constraints of preious work has resulted in an underestimation of the contribution of these species to global catch and revenue, and that the proportoin of transboundary species is larger than previouslly estimated.

We overlaid the known distribution of 938 commercially valuable marine species responsable for an average of 96.5% of global EEZ catch reconstruced by the *Sea Around Us* at the species level between 2005 and 2014, and the EEZ of 153 coastal countries. We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level. Also, recent research has suggested strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis. For this analysis we only considered shared species between neighboring EEZs, rather than the species' extended distribution (e.g. we did not include the high seas). We relay in multiple data sources including occurance, distribution models and catch data and only considered a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods). Finally, each neighbouring EEZ had to have more than 25% of a species' distribution in order for the species to be categorized as transboundary.
<!-- ## Results and discussion (~1000 words) -->

We identified 633 transboundary exploited species in the world, more than double of previous estimates [@Teh:2015gd], representing 67.5% of the species analyzed. Between 2005 and 2014, national fleets targeting transboundary species within EEZs landed an average of 47.6 million tonnes, representing 79.9% of the EEZ-derrived catches reconstructed by the *Sea Around us* at the species level, and generating USD 65,483 million in fishing revenue. This is equivalent to 78.5% of the global fishing revenue of the same period. These values are more than the 34.2 million t and USD 30,700 million in fishing revenue previously estimated [@Teh:2015gd]. Based on our assesment (see Methods), 590 transboundary species were also identified as discrete species. This is the case of species such as Peruvian anchoveta (*Engraulis ringens*) that is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile. The contribution of discrete species (e.g. the anchoveta catch of Ecuador) was 14.8 million tonnes representing USD 22,167 million in fishing revenue. Within these species, 305 species were considered exclusiveley discrete species, that is, they could be fished by neighbouring nations but were not considered transboundary. These species represented only 2.7 million tonnes and 3,691 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level (Fig. 1).

```{r Call_Map, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary marine species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by colour coding of EEZs and land polygons, respectively. B) Contribution of tranbsoundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD"}

@@

![Fig. 1. Number of transboundary marine species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by colour coding of EEZs and land polygons, respectively. B) Contribution of tranbsoundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD.](./Figures/Fig1.png)

We aggregate countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *DF*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). China (USD 6,279 million) in E. Asia and the US (USD 9,605 million) in N. America contribute 55% and 83% to their respective region's revenue, respectively. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 30% of the fishing global revenue from transboundary species. Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead transboundary catch statistics harvesting over 5,000 thousand tones each on average between 2005 and 2014 (Fig S1). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and accounting for 79% of the transboundary species production. Similarly, Russia shares the Alaskan polllock (*Theragra chalcogramma*) fishery with both Japan and the US. Despite the large catch volume of Peru and Russia, regional catches are only significantly different in E. Asia (one-way ANOVA; *Df*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons harvested in 2016 [@FAO:2018KK] getting around 1/3 of its capture from transboundary species, followed by Japan (4.2 million tonnes) and South Korea (2.3 million tonnes). Differences in tansboundary fisheries can also be ilustrated in terms of catch-revenue/EEZ $km^2$ (Figure S2). As an example, the aggregated area of all EEZs of Northern Europe is 3.3 million $Km^2$, the $5^{th}$ smallest of the 17 groups analized (Table S1), however, it has the most revenue (USD 246.7 million per $Km^2$) and catch (260.1 tonnes per $Km^2$) of transboundary species per EEZ area, but only significantly differing from Sub-Saharan African and Southern Europe countries (*p* < 0.05).

We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 2) [@Kleisner:2011wn]. Previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, over-exploited or collapsed [@Kleisner:2011wn]. Following this concept, Category A would represent fisheries that are re-building or under-exploited, category B would be species that are sustainably harvested, and Category C would be species that are over-exploited. We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,446) = 53.9, *p* < 0.001). In the case of discrete species, only Category C is significantly different than the rest (one-way ANOVA, *Df*(2,446) = 52.88, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.001). On the other hand, for transboundary species, all categories are significantly different (one-way ANOVA, *Df*(2,446) = 52.88, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05 for all categories). While many environmental and social-economic factors (e.g. temperature, markets, fishing policies) affect fishing effort and catch volume [@Pauly:2013gs; @Branch:2008kk], this result suggests that the management of discrete species could be better off as the variation between categories A and B are closer (*p* = 0.052) than for transboundary species (*p* = 0.002).

@@

Empirical analysis suggests that in most cases, management of transboundary species will have better outcomes in terms of fish yields when nations cooperate [@Bailey:2010ga]. Examples of successful co-management include Canada and the United States over Pacific salmon [@Song:2017va], Norway and Russia over Atlantic cod (*Gadus morhua*), haddock (*Melanogrammus aeglefinus*) and capelin (*Mallotus villosus*) in the Barents Sea [@Diekert:2010gp; @Eide:2013dha], and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration over shared stocks may threaten stock sustainability [@Clark:1980js; @Nguyen:2018dk], and reduce the future profitability potential of the fishery [@Merino:2007jz]. Yet, we acknowledge that cooperation or lack thereof can be a complex process [@Jensen:2015cf] and that there are specific cases where cooperation might not be the best strategy [@Munro:2015df]. Collaboration can also reduce conflict between nations over shared resources as climate change shifts the distribution of species creating tension between countries. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations were introduced to compensate for losses [@Munro:2002uf]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for Atlantic mackerel (*Scomber scombrus*) [@Spijkers:2017ij], a species that is transboundary and contributed an average of 598.4 thousand tonnes to catches (8.19% of total) and USD 850.2 million in revenue (7.17% of total) to the countries involved in the dispute between 2005 and 2014. The relatively small contribution of mackrell to the involved countries suggests that even a small fraction of the year's catch can lead to important conflicts between neighboring countries. In addition to past and present conflicts, climate change is likely to exacerbate conflicts over transboundary species due to shifts in their distribution, requiring flexible agreements that accommodate future uncertainties [@Pinsky:2018cb].

<!-- **NOTE: DOUBLE CHECK REFERENCES eddie and derek]** -->

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species lke greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,109) = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 3).

@@

Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory pelagic fish species are managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species [@CullisSuzuki:2010fi]. We estimated that 81.8% of these species are shared across neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighboring RFMOs (Fig. 3). The effectiveness of RFMOs responsible for managing high migratory stocks has been previously questioned [@CullisSuzuki:2010fi; @Lodge:2007bt]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 13 species (40 stocks) suggests that most tuna and billfish stocks are sustainably harvested with only three stocks currently considered as overfished [@Pons:2016ep]. In the other hand, species with narrower ranges are often managed at the national level due to localized populations, however, the potential boundary-crossing of these species must not be ignored [@Payne:2008vm]. The world of fisheries is highly connected through larvae dispersal [@Ramesh:2019va] and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. In addition, coral regions often share multiple species, making coordination among countries key to achieve fisheries and conservation goals [@Levin:2018in; @Payne:2008vm]. Many coastal communities, particularly in developing regions, are highly dependent on subsistance fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. 3B) are among the highest fish consuption countries (> 10 g of fish protein per capita per day) [@FAO:2018KK]. In these countries, both pelagic and reeef fishes contribute substantially to both local food security and livelehood [@Cabral:2018hc]. In the case of Kiribati, the Solomon Islands and Papua New Guinea, they are parties to the Pacific Nauru Agreement that sustainably manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

In our analyses we did not seek to include climate change as a driver of species’ distribution. Distribution shifts are the most commonly recorded response of marine species to the effects of climate change - chiefly warming - in the ocean [@Poloczanska:2016kk]. Catch composition might have already changed in the last five decades [@Cheung:2013gk] and will continue throuugh the $21^{st}$ century [@Cheung:2016jd]. Previous study projected that climate change will bring at least one new transboundary species per EEZ in the future [@Pinsky:2018cb], while also change the proportion of the distribution (e.g. Area Index) that each county’s share to a degree that could influence fisheries management (*Palacios-Abrantes et.al. under review*). Exploring the detailed impacts of climate change in share species is key to develop local adaptation methods to minimize the effect of such distributional change on the conservation of species as well as the economy and food security of coastal nations.

Our study identified species currently shared by neighboring coastal nations and highlights the importance of these species' contribution to global capture and revenue from wild fisheries. Our results show that captures and revenues from shared species are substantially higher than those from discrete species, as well as than previously estimated. This result highlights the importance of transboundary fisheries and their potential contribution to food security and wellbeing. Moreover, we show significant differences in the management of transboundary and discrete species, suggesting a need for improving the management of transboundary fisheries. Previous work has found that collaboration will result in better outcomes for shared marine resources [@Bailey:2010ga]. Thus, countries should pay specific attention in the management of the identified transboundary stocks potentially creating collaborative management plans. Identifying existing transboundary species is the first step towards joint management frameworks that are precautionary, strive for sustainability, and can be flexible to accommodate the uncertain future driven by climate change.

<!-- # **Acknowledgments* * -->

<!-- ## Funding: include complete funding information;  -->
<!-- ## Authors contributions: a complete list of contributions to the paper (we encourage you to follow the CRediT model), -->
<!-- ## Competing interests: competing interests of any of the authors must be listed (all authors must also fill out the Conflict of Interest form). Where authors have no competing interests, this should also be declared. -->

## Data and materials availability

All data is available from the authors upon request.

# Materials and Methods

To ascertain whether or not a species can be considered transboundary, we developed a method based on a series of criteria and the overlay of species' distributions and the spatial boundaries of the world’s EEZs. All analyses were conducted at the species level and we consider transboundary species as those that occur within the Exclusive Economic Zone (EEZ) of neighboring countries [@Munro:2002uf]. All data was scaled to $0.5^o$ x $0.5^o$ lat-long grid cells.

@@

To determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs we used the reconstructed list of commercial marine fish and invertebrate species from the *Sea Around Us* (SAU) database (http://www.seaaroundus.org) and determined their current distributions based on an average from 1970-2000. The total number of species analized was 938 representing 54% of the average catch reported by the *Sea Around Us* between 2005 and 2014, including generalized categories such as *Marine fish not identified* or *Mollusca*. However, because we carried the analysis at the species level, we did not include such generalized categories in our comparrisons. Our final sample represents 63.2% of identified species, accounting for 96.52% of the catch identified at the species' level. To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table 1). Each source represents a different method of estimating the distribution of a given species and thus, provides a more robust result. Only commercial fished species with data from all four sources were included in the analysis.

>>>>

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
# subtitle: "[DRAFT DO NOT SHARE]"
subtitle: "Most neighbor nations share marine exploited species"
output:
  word_document:
    reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ./References/FishForVisa_Ref.bib
csl: ./References/science.csl
---

* **Corresponding author**: j.palacios@oceans.ubc.ca

@@

# **One Sentence Summary**
<!-- *A brief teaser statement highlighting main result of the paper, understandable by a scientist not in your field, without jargon or abbreviations. This will appear online adjacent to the title and should not repeat phrases already present there* -->

@@

# **Abstract**
<!-- *An opening sentence that sets the question that you address and is comprehensible to the general reader, background content specific to this study, results, and a concluding sentence. It should be a single paragraph.* -->

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 80% of the global-EEZs catches (48 million tons) and 78.5% of fisheries revenue (USD 65 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. This study highlights the large and under-recognized challenge in achieving international fisheries management goals, particularly as species redistribute under climate change.

# **Main Article**
<!-- # Introduction (~555/500 words) -->

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors defined by their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. Thus, a species can have a broader distribution (cosmopolitan) or be endemic [@Nelson:2016kk]. However, fisheries management is predicated on the definition of “stocks” that, in most cases, are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, established political boundaries across the distribution of many species, creating shared species between nations. Theories and empirical experience have shown that fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@Munro:2002uf]. For instance, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and for fisheries management [@Miller:2013iv].

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that are shared by neighboring nations (hereafter called transboundary species [@Munro:2002uf]) is still unclear. Early data limitations on species' distributions allowed only for an informed guess on the number of transboundary species [@Caddy:1997ue] while a more recent study explored the historical contribution of shared stocks following a literature review [@Teh:2015gd]. Thus, no method has included the actual distribution of species in the quantification of transboundary stocks. Here, we aimed to estimate the number of exploited marine species shared by neighboring countries and determine their contribution to global and regional catch and fisheries revenue. Moreover, we categorize species according to their exploitation rate while identifying differences by habitat preference. We hypothesize that the methodological constraints of previous work has resulted in an underestimation of the contribution of these species to global catch and revenue, and that the proportion of transboundary species is larger than previously estimated.

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 241 EEZs of 198 coastal territories. We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. We rely on multiple data sources including occurrence, distribution models and catch data and only considered a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

<!-- ## Results and discussion (~1000 words) -->

We identified 633 transboundary exploited species in the world, more than double the previous estimates [@Teh:2015gd], representing 67.5% of the species analyzed. Between 2005 and 2014, national fleets targeting transboundary species within EEZs cached an average of 47.6 million tons, representing 79.9% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating USD 65,483 million in fishing revenue. This is equivalent to 78.5% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 30,700 million in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. Based on our assessment (see Methods), 590 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile, and thus discrete for Ecuador. The contribution of discrete species was 14.8 million tons representing USD 22,167 million in fishing revenue. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary. These species represented only 2.8 million tons and 3,691 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level (Fig. 1).

```{r Call_Map, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD"}

@@

![Fig. 1. Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD.](./Figures/Fig1.png)

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (mean revenue = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ 2,667 million USD). The estimated revenues from transboundary species in these two regions are significantly different from other regions in terms of revenue (one-way ANOVA; *DF*(16,158) = 4.948, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). China (USD 6,279 million) in E. Asia and the USA (USD 9,605 million) contribute 55% and 82% to the total revenues from the E. Asia and N. America regions, respectively. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,368) are amongst the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 40% of the global fisheries revenue from transboundary species. Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead transboundary catch statistics harvesting over 5.8 million tons annually each on average between 2005 and 2014 (Fig S1). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and account for 79% of the transboundary species production. Similarly, Russia shares the Alaskan pollock (*Theragra chalcogramma*) fishery with both Japan and the USA. Transboundary species also contribute largely to fisheries in E. Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). Particularly, China the world's top fish producer [@FAO:2018KK] getting one third (5.1 million tons) of the total catches from transboundary species, followed by Japan (1.8 million tons) and South Korea (1.06 million tons). However, when we account for the total area of waters in different EEZs and regions relative to their catches, the importance of transboundary species to European fisheries become apparent (Fig. S2, Table S1).

We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 2). Previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, over-exploited or collapsed [@Kleisner:2011wn]. We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,447) = 27.091, *p* < 0.001). Specifically, transboundary species that are only fished by one country are generally less likely to be over-exploited compared to those shared by neighboring countries (Table S2). Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,447) = 55.01, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,380) = 21.86, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.052). This result suggests that the management of discrete species could be better off as the variation between Categories A and B is closer (*p* = 0.052) than for transboundary species (*p* = 0.001).

@@

Empirical analysis suggests that cooperation in the management of transboundary stocks will have better outcomes, although there are specific cases where cooperation might not be the best strategy [@Munro:2002uf]. Examples of successful cooperative-management include Norway and Russia over Atlantic cod (*Gadus morhua*) [@Eide:2013dha] and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration over shared stocks may threaten stock sustainability, reduce the future profitability potential of the fishery, and create conflict between neighboring nations [@Clark:1980js; @Spijkers:2017ij]. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Mackerel is a transboundary species that contributed an average of 598.4 thousand tons to catches (8.19% of total) and USD 850.2 million in revenue (7.17% of total) to the countries involved in the dispute between 2005 and 2014. The relatively small contribution of mackerel to the involved countries suggests that even a small fraction of the year's catch can lead to important conflicts between neighboring countries. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed impacts of climate change in share species is key to develop local adaptation methods to minimize the effect of such distributional change on the conservation of species as well as the economy and food security of coastal nations.

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 3).

@@

Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species [@CullisSuzuki:2010fi]. We estimated that 81.8% of these species are shared across neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighboring RFMOs (Fig. S3). The effectiveness of RFMOs has been previously questioned [@CullisSuzuki:2010fi]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 40 tuna and billfish stocks suggests that most stocks are sustainable harvested with only three currently considered as overfished [@Pons:2016ep]. In the other hand, species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm]. However, the potential boundary-crossing of these species must not be ignored. The world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. In addition, coral regions often share multiple species, making coordination among countries key to achieve fisheries and conservation goals [@Payne:2008vm]. Many coastal communities, particularly in developing regions, are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. 3B) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc]. In the case of Kiribati, the Solomon Islands and Papua New Guinea, they are parties to the Pacific Nauru Agreement that sustainable manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

Our study identified species currently shared by neighboring coastal nations and highlights the importance of these species' contribution to global capture and revenue from wild fisheries. Our results show that captures and revenues from shared species are substantially higher than those from discrete species, as well as than previously estimated. This result highlights the importance of transboundary fisheries and their potential contribution to food security and wellbeing. Moreover, we show significant differences in the management of transboundary and discrete species, suggesting a need for improving the management of transboundary fisheries. Previous work has found that collaboration will result in better outcomes for shared marine resources [@Munro:2002uf]. Thus, countries should pay specific attention in the management of the identified transboundary stocks potentially creating collaborative management plans. Identifying existing transboundary species is the first step towards joint management frameworks that are precautionary, strive for sustainability, and can be flexible to accommodate the uncertain future driven by climate change.

# **References**

<div id="refs"></div>

\clearpage

# **Acknowledgments**

We thank T. Clark and A. Tidd for helpful comments and discussions. **Funding**: G.R. was supported by the EO Wilson foundation. C.W. was supported by the Walton Family Foundation (grant 2018-1371), the David and Lucile Packard Foundation (grant 2019-68336), and the Gordon and Betty Moore Foundation (grant GBMF5668.02). W.W.L.C was supported by the NSERC Discovery Grant and the SSHRC Partnership. **Authors contributions**: J.P.A. Conceptualization, methodology, formal analysis, writing original draft. G.R. Methodology, data curation, writing - review & editing. C.W. Methodology, writing - review & editing. W.W.L.C. Supervision, writing - review & editing. All authors provided valuable input to the study conceptualization and discussed the results and analysis. **Competing interests**: No declared. **Data and materials availability**: The Sea Around Us datasets used are publicly available at http://www.seaaroundus.org/, the occurrence dataset is all publicly available (see Methods), the ENM-Nereus data can be accessed upon request to G.R. All code is available at https://github.com/jepa/FishForVisa

## Data and materials availability
All data is publicly available online and upon request from the authors.

## List of Supplementary Materials
Figures S1-S4
Tables S1 -S2
References in main manuscript (1-27)

# **Materials and Methods**

@@

To determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs we used the reconstructed list of commercial marine fish and invertebrate species from the *Sea Around Us* (SAU) database (http://www.seaaroundus.org) and determined their current distributions based on an average from 1970-2000. The total number of species analyzed was 938 representing 54% of the average catch reported by the *Sea Around Us* between 2005 and 2014, including generalized categories such as *Marine fish not identified* or *Mollusca*. However, because we carried the analysis at the species level, we did not include such generalized categories in our comparisons. Our final sample represents 63.2% of identified species, accounting for 96.52% of the catch identified at the species' level. To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table 1). Each source represents a different method of estimating the distribution of a given species and thus, provides a more robust result. Only commercial fished species with data from all four sources were included in the analysis.
